---
title: "Visualizing Starwars characters"
author: "Aurelia Ponzio"
format: html
editor: visual
---

```{r}
#| label: load-packages
#| include: false
library(tidyverse)
library(ggridges)
```

1.  Glimpse at the starwars data frame.

```{r}
#| label: glimpse-starwars
glimpse(starwars)
```

2.  Modify the following plot to change the color of all points to `"pink"`.

```{r}
#| label: scatterplot
ggplot(starwars, 
       aes(x = height, y = mass, color = gender, size = birth_year)) +
  geom_point(color = "pink")
```

3.  Add labels for title, x and y axes, and size of points. Uncomment to see the effect.

```{r}
#| label: scatterplot-labels
ggplot(starwars, 
       aes(x = height, y = mass, color = gender, size = birth_year)) +
  geom_point(color = "#30509C") +
  labs(
    title = "Mass versus Height of Star Wars Characters",
    x = "Height (cm)", 
    y = "Weight (kg)",
    size = "Year Born (BBY - Before Battle of Yavin)"
    )
```

4.  Pick a single numerical variable and make a histogram of it. Select a reasonable binwidth for it.

(A little bit of starter code is provided below, and the code chunk is set to not be evaluated with `eval: false` because the current code in there is not valid code and hence the document wouldn't knit. Once you replace the code with valid code, set the chunk option to `eval: true`, or remove the `eval` option altogether since it's set to `true` by default.)

```{r}
#| label: histogram
#| eval: true
ggplot(starwars, aes(x = -birth_year)) +
  geom_histogram(binwidth = 15)
```

5.  Pick a numerical variable and a categorical variable and make a visualization (you pick the type!) to visualization the relationship between the two variables. Along with your code and output, provide an interpretation of the visualization.

```{r}
#| label: num-cat
ggplot(starwars, aes(x = -birth_year, y = sex, color = sex, fill = sex)) +
  geom_density_ridges(alpha = 0.5)
```
By creating a ridgeline plot of birth year by sex, we overwhelmingly see that
the vast majority of characters were born around 25 BBY. 
We see a very interesting double spike in none, as a result of droids.
That one data point pre-750 is likely Yoda.

6.  Pick a single categorical variable from the data set and make a bar plot of its distribution.

```{r}
#| label: barplot
ggplot(starwars, aes(x = sex)) + 
  geom_bar()
```

7.  Pick two categorical variables and make a visualization to visualize the relationship between the two variables. Along with your code and output, provide an interpretation of the visualization.

```{r}
#| label: cat-cat
ggplot(starwars) + 
  geom_bar(aes(y = sex)) +
  facet_wrap(. ~ species)
```

It's a bit difficult to view this chart, but one can see that the vast
majority of species only have males. Of species with females, most other than humans are fairly equal. We can see here the humans are the largest species, followed by droids, followed by Gungans.

8.  Pick two numerical variables and two categorical variables and make a visualization that incorporates all of them and provide an interpretation with your answer.

(This time no starter code is provided, you're on your own!)

```{r}
#| label: multi
subset(starwars, name != "Jabba Desilijic Tiure") %>%
  drop_na(sex) %>%
  ggplot() +
    geom_point(aes(x = height, y = mass, color = birth_year)) +
    facet_grid(. ~ sex)
```

We can see that interestingly droids come in three sets. Most females are around the same weight/mass, with one outlier. Males have a much larger spectrum.