---
title: "Airbnb listings in Edinburgh"
author: "Aurelia Ponzio"
date: "2023-01-19"
output: html_document
---

This work is based on the questions found at https://datasciencebox.org/course-materials/hw-instructions/hw-02/hw-02-airbnb-edi.html

## Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse)
library(dsbox)
library(ggridges)
```

## Exercises

### Exercise 1

The dataset has the number of rows displayed below
```{r}
nrow(edibnb)
```

### Exercise 2

Each row found by running `View(edibnb)` represents an individual air bnb listing.

### Exercise 3

I've wrapped the histogram because it would take up far too much space if 
they were stacked in a column.

```{r prices-neighbourhoods}
ggplot(data = edibnb, mapping = aes(x = price)) +
  geom_histogram(binwidth = 25) +
  facet_wrap(~neighbourhood)
```

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r top-5-median}
top5 = edibnb %>%
  filter(!is.na(price)) %>%
  group_by(neighbourhood) %>% 
  summarise(med_price = median(price)) %>%
  top_n(n = 5, med_price) %>%
  arrange(desc(med_price)) %>%
  print()


  #ggplot(data = edibnb, mapping = aes(x = price, y = neighbourhood, fill = neighbourhood, color = top5$neighbourhood)) +
   # geom_density_ridges(alpha = 0.5)
```

```{r top-5-median-summary-stats}
# remove this comment and add the code for Exercise 4 here
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r review-scores-rating}
# remove this comment and add the code for Exercise 5 here
```