---
title: "Data Transformation Readings"
author: "Aurelia Ponzio"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This assignment included chapter 5 of R for Data Science. 
The reading can be found at https://r4ds.had.co.nz/transform.html

As always, we begin with tidyverse.
We'll use nycflights13 for our dataset.
``` {r}
#| label: load-packages
#| warning: false
#| message: false
library(tidyverse)
library(nycflights13)
```


## 5.2.4 Finding rows with filter

### 1. Find all flights that
  Had an arrival delay of two or more hours
  Flew to Houston (IAH or HOU)
  Were operated by United, American, or Delta
  Departed in summer (July, August, and September)
  Arrived more than two hours late, but didn’t leave late
  Were delayed by at least an hour, but made up over 30 minutes in flight
  Departed between midnight and 6am (inclusive)

``` {r}
# arrival delay >= 2 hours
flights %>%
  filter(arr_delay >= 120)

# flew to Houston
flights %>%
  filter(dest == "IAH" | dest == "HOU")

# united, american, or delta
flights %>%
  filter(carrier == "UA" | carrier == "AA" | carrier == "DL")

# departed in summer
flights %>%
  filter(month == 7 | month == 8 | month == 9)

# more than two hours late when on time
flights %>%
  filter(dep_delay <= 0 & arr_delay > 120)

# delayed by an hour but made up over 30 mins
flights %>%
  filter(dep_delay >= 60 & arr_delay <= 30)

# departed between midnight and six am
flights %>%
  filter(hour >= 0 & hour <= 6)
```

### 2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?
According to the docs it is a way to find if a numeric value falls in a range.
It can be used to simplify the following questions:
``` {r}
# departed in summer
flights %>%
  filter(between(month, 7, 9))

# departed between midnight and six am
flights %>%
  filter(between(hour, 0, 6))
```

### 3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
According to the below code, 8255 flights are missing their departure time.
These are flights that never took off.
``` {r}
flights %>%
  filter(is.na(dep_time)) %>%
  count()
```

Looking at the data frame, other missing variables include dep_delay, 
arr_time, arr_delay, tailnum, and air_time. 
Other than air_time, all of these can be accounted for by the plane either
not taking off or not landing.
Missing tailnums are probably flights that were registered but never given planes.

### 4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)
NA ^ 0 isn't missing because R assumes that any number to the 0th is 1.
NA | TRUE isn't missing because the the second option is true so it will evaluate to true regardless of the value for NA.
FALSE & NA isn't missing for the same reason as above but flipped. Anything & FALSE will
evaluate to false.
The general rule is that NA stands in for some unknown so if the expression can be evaluated without that unknown it will be.


## 5.3.1 Arrange rows with Arrange

### 1. How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
``` {r}
flights %>%
   arrange(!is.na(flights))
```

### 2. Sort flights to find the most delayed flights. Find the flights that left earliest.
``` {r}
# earliest 
flights %>%
  arrange(dep_delay)

# most delayed
flights %>%
  arrange(desc(dep_delay))
```

### 3. Sort flights to find the fastest (highest speed) flights.
``` {r}
flights %>%
  arrange(distance / air_time)
```

### 4. Which flights travelled the farthest? Which travelled the shortest?
``` {r}
# farthest
flights %>%
  arrange(distance)

#shortest
flights %>%  
  arrange(desc(distance))
```

## 5.4.1 Select columns with select()
### Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
List them explicitly, list everything else explicitly, create a regex that matches,
use starts_with or ends_with.

### What happens if you include the name of a variable multiple times in a select() call?
``` {r}
flights %>%
  select(month, day, month, day)
```
It ignores the second call.

### What does the any_of() function do? Why might it be helpful in conjunction with this vector?
``` {r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
flights %>%
  select(any_of(vars))
```
Lets you provide an easy list and select based on it.

### Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?
``` {r}
select(flights, contains("TIME"))
```
I am a bit suprised it ignores case by default
This can be changed with `ignore.case = FALSE`
``` {r}
select(flights, contains("TIME", ignore.case = FALSE))
```

## 5.5.2 Adding new variables with mutate

### 1.Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
``` {r}
mutate(flights,
  dep_time,
  hour = dep_time %/% 100,
  minute = dep_time %% 100
)
```

### 2. Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?
Air time is wheels up to wheels down, departure time to arrival time is gate to gate.
This can't really be "fixed" but it's something to remain aware of.
We can calculate taxi time if we want.

### 3. Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?
Departure time minus scheduled depature time should equal depature delay.

### 4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().
``` {r}
flights %>% 
  mutate(rank = min_rank(-dep_delay)) %>%
  select(dep_delay, rank) %>%
  arrange(rank)
```


### 5. What does 1:3 + 1:10 return? Why?
``` {}
1:3 + 1:10
```
```
Warning message:
In 1:3 + 1:10 :
  longer object length is not a multiple of shorter object length
```

This gives an error, because the two different frames are seperate lengths.
It doesn't know how to add them.

### 6. What trigonometric functions does R provide?
The following:
```
cos(x)
sin(x)
tan(x)

acos(x)
asin(x)
atan(x)
atan2(y, x)

cospi(x)
sinpi(x)
tanpi(x)
```

## 5.6.7 Grouped summaries with summarise()

### Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights.
Group by arrival delay, departure delay, and average of the two. 
Take median/mode of each

### Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).
We can use sum(!is.na(x))

### Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?
No departure delay only is the best definition, 
because it's possible for flights to have a depature delay but no arrival 
delay and not have been cancelled due to a crash.

### Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?
``` {r}
flights %>%
  group_by(month, day) %>%
```