---
title: "Putney Alumni"
author: "Aurelia Ponzio"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Time to do some web scraping!

Going to start with a list of notable alumni found on Putney's website.

First, we have to check if we're allowed.

``` {r, warning = FALSE}

library(robotstxt)
paths_allowed("https://en.wikipedia.org")
```

There we are! Time to get going!

Let's add our libraries.
``` {r, echo = FALSE, warning = FALSE, message = FALSE}
library(tidyverse)
library(rvest)
library(stringr)
```

For this, I'll be using the page https://en.wikipedia.org/wiki/Category:The_Putney_School_alumni
to get a list of alumni. This information will include their name (string), year of graduation (integer), occupation (string) and a link to their Wikipedia page (string).

This will not be a particularly interesting data set but it's a nice simple example for me to get comfortable with the tools

First, we read the whole page and then we convert it to a nicer form
```{r}
page <- read_html("https://en.wikipedia.org/wiki/Category:The_Putney_School_alumni")
class(page)
```

Next, we extract each person's name as well as a link to their Wikipedia page and store it in a tibble.

``` {r}
names <- page %>%
  html_nodes(".mw-category-group a") %>%
  html_text()

links <- page %>%
  html_nodes(".mw-category-group a") %>%
  html_attr("href")

putney_alums <- tibble(
  name = names,
  link = links
)

putney_alums <- mutate(putney_alums, link = paste0("https://en.wikipedia.org", putney_alums$link))

putney_alums

```
