---
title: "Reading for January 19"
author: "Aurelia Ponzio"
date: "2023-01-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This assigment included reading 3.6 in 
https://r4ds.had.co.nz/data-visualisation.html

The exercises are from 3.6.1

We'll begin by using tidyverse.

``` {r}
#| label: load-packages
#| warning: false
#| message: false
library(tidyverse)
```

#### 1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
For a line chart, one would use  `geom_line()`. For a boxplot, `geom_boxplot()`,
for a histogram `geom_histogram()`, and for an area chart `geom_area()`.

#### 2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
``` {r}
#| message: false
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

I expect this to create a plot using data from the mpg dataset.
It will create a scatter plot and a smooth line fitted to the data.
The plot will have `displ` on the x axis, `hwy` on the y axis, 
and will be colored according to `drv`.
The smooth plot will have no confidence intervals.

After running it, it looks how I expected. I was initially unsure how 
`geom_smooth()` would handle color but three lines makes perfect sense.

#### 3. What does `show.legend = FALSE` do? What happens if you remove it? Why do you think I used it earlier in the chapter?
`show.legend = FALSE` removes the legend, 
the section explaining what each color/symbol means.
It was used earlier so all three plots would fit nicely.

#### 4. What does the `se` argument to `geom_smooth()` do?
If true, it shows confidence intervals. If false, it doesn't.

#### 5. Will these two graphs look different? Why/why not?
``` {r}
#| message: false
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

They'll look the same. The first is just the second but written better.

After viewing them, I was correct

#### 6. Recreate the R code necessary to generate the following graphs.
``` {r}
#| message: false
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(aes(group = drv), se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(aes(color = drv)) + 
  geom_smooth(se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(aes(color = drv)) + 
  geom_smooth(se = FALSE, aes(linetype = drv))

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(stroke = 4, color = "white") +
  geom_point(stroke = 2, aes(color = drv))
```
